|

### 2020-10-16 documents/ovando-etal-assessing-global-fisheries.Rmd
pre submission draft

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess. Obtaining accurate estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of high-priority data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
---

@@

rand_fits <- assess_ram_fits %>% 
  filter(data == "ram-data") %>% 
  mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
         data = "guess")

perf <- assess_ram_fits %>% 
  bind_rows(rand_fits) %>% 
  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
  group_by(data) %>% 
  summarise(
    mpe = median((mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
    mape = median(abs(mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
  ) %>% 
  arrange((mape))

@@

<!-- * more assertive on what the data needs really are -->
<!--   * where to apply globally  -->
<!-- * For places with less coverage, what would we actually recommend there  -->
<!-- * what do we know about catchabiliy -->
<!-- - e.g. 51, fleets became industrial  -->
<!--   - hammer down on that's the issue with the effort  -->
<!--   - how do we translate effort into f bey area -->
<!--   - here's  -->

# Publication Requirements

Target Journal: [Nature Sustainability](https://www.nature.com/natsustain/about/content)

3500 words main text

50 references (main text)

3000 words (max) methods

@@

The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". A substantial part of achieving this goal depends on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems. Meeting these targets depends in part on our ability to effectively track the status of global marine fish stocks. The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking the global state of fisheries. As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020]. While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment. Although revised periodically over time, the method has kept revisions to minimally necessary levels for the purposes of temporal consistency and comparability. With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.

49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion. While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable (and expert opinion can also be misinformed). Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018;@thorson2012a].

Formal stock assessments require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist [@maureaud2020], and would be prohibitively expensive to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment. Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013]. In addition, a global assessment requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time. As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks. Each of the published catch-based methods reviewed in Free et al. 2020 [@free2020] sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020]. A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.

Examining stock assessment data from RAM from the  Northeast Pacific,the Northeast Atlantic and the Mediterranean & Black Sea, the Northeast Pacific is in the best state (mean RAM B/B~MSY~ = 1.61), NE Atlantic (median RAM b/b~MSY~ =1.27) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.52) [@hilborn2020]. The Costello et al. 2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or NOrtheast Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch and equates these to stock status, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)). While the methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks,

@@

readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fisheries? We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability. However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management through Fisheries Management Index (FMI) scores [@melnychuk2017], trawl-fishing footprints through swept-area-ratio (SAR) values [@amoroso2018], and life history [@thorson2017c]. `sraplus` provides a flexible framework for integrating these data into valid statistical model.

The coming decades are a critical period for the future of global oceans. Climate change, population growth, and economic expansion will all place increasing pressure on fisheries. While our understanding of global fisheries has improved greatly,  mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world, particularly those stocks that are currently unassessed. At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods. We leverage these opportunities to provide a global assessment of our perception of global stock status resulting from different use of different data sources. Our results demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection and interpretation effort.

@@

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data (combinations of catches, effort, fisheries management index, and swept area ratio values) relative values reported in the RAM database.

@@

For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores. We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock. We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so. We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status. We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`. It is critical to note that in both of these cases the catch data on its own contains no information about stock status; we can only produce estimates of stock status by combining catch data with explicit priors and specified functional relationships [@thorson2013]. This is important since it means that we must look to our priors and functional forms to understand where our insight stock status is coming from, not the catch data alone. Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.

Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one. CMSY, using no stock-specific information but the catch history and the estimated resilience of the target species, badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing. Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~. Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit. This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values. An abundance index based on catch and regional effort cannot capture this trend, since it interprets the extremely low catches (near zero) as low biomass since we do not have sufficiently fine-scale effort data to reflect the reduction in effort that goes hand-in-hand with regulated catch reductions (Fig.\@ref(fig:cs-plot)).

```{r cs-plot, fig.cap="Estimates of stock status for case study fisheries using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index and Swept Area Ratio scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%. Each point is a stock, point size is a function of stock size. The shaded hexagons in the background show the distribution of stock status estimates from the RAM Legacy Stock Assessment Database for each of the stocks included in this figure, with darker colors representing increased concentrations of RAM stocks."}

ex_kobe_plot
```

That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise. In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process. Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.

While SAR and FMI data are available for a wide number of stocks, they are not ubiquitous enough to provide stock-specific priors for the number of fisheries needed for a global assessment. In addition, predicting stock status based on SAR, FMI, or catch patterns requires that the same patterns that exist within the training data as exist in other stocks to be used for prediction, which is far from clear given the highly managed nature of the stocks for which we have reliable estimates of stock status. We can extend the ideas presented in the case study, using combinations of catch, SAR, FMI, and effort data to produce estimates of stock status at the global level.

As our test case we use `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM with estimates of B/B~MSY~ and greater than 25 years of continuous catch history. B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then pair the catch histories for these stocks with regional-level swept area ratio, fisheries management index, and effort data. For any given stock we use the area-weighted swept area ratio (where area refers to the spatial area of the SAR study in question) for the FAO region of the stock. For fisheries management index we use the mean component FMI values for the species group and FAO region in question (or FAO region if no matches exist at the species group level). This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries. We then constructed an effort index for the stock in question as the total effort from @rousseau2019 estimated in all countries that report catch of the species in question within its FAO region, weighting the effort of each country by its contribution to the total catch of that species in that FAO region.

As a proof of concept, we first estimated B/B~MSY~ of our candidate RAM stocks by using `sraplus` to fit to an abundance index drawn directly from RAM. We then fit a range of models utilizing different combinations of the regional-level data made available to them (Table.\@ref(tab:dat-desc)). We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy is calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).

```{r dat-desc}
dat_description <- tribble(~`Data Name`, ~Description,
        "RAM-Data", "Fit to abundance index from RAM",
        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.5% tech. creep",
        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.5% tech. creep",
        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
        "catch-only", "priors on terminal B/Bmsy informed by catch history",
        "heuristic", "priors on initial and terminal B/Bmsy informed by Cmsy heuristc")

Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`. This creates two main reasons why `sraplus` might produce poor results at a global level: model misspecification and poor data quality. Model misspecification errors might result from the surplus production model underlying `sraplus` being fundamentally unable to capture the more complicated dynamics resulting from the full age-structured population models common in RAM. Errors resulting from poor quality data mean that even if the underlying model is a reasonable approximation of the real world, the model may produce incorrect results due to the data supplied to the model simply not containing accurate information on the phenomenon in question.

Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality. Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

<!-- The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias. The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time. Looking across space, we see clear geographic patterns in model performance. The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada). The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.  -->

>>>>

---
title: |
  | Global Assessments of Fishery Status need 
  | Better Data more than Better Models
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::pdf_document2: default
  bookdown::word_document2:
    reference_docx: word-template.docx
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 49% of global catch, our knowledge of the state of the remaining 51%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fisheries remains poor, even when new sources of broadly available data are added. Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data at around the globe, rather than the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
---

@@

# rand_fits <- assess_ram_fits %>%
#   filter(data == "ram-data") %>%
#   mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
#          data = "guess")

perf <- assess_ram_fits %>% 
  # bind_rows(rand_fits) %>% 
  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
  group_by(data) %>% 
  summarise(
    mpe = median((mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
    mape = median(abs(mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
  ) %>% 
  arrange((mape)) %>% 
  rename("Data Used" = "data")

@@

<!-- * more assertive on what the data needs really are -->

<!--   * where to apply globally  -->

<!-- * For places with less coverage, what would we actually recommend there  -->

<!-- * what do we know about catchabiliy -->

<!-- - e.g. 51, fleets became industrial  -->

<!--   - hammer down on that's the issue with the effort  -->

<!--   - how do we translate effort into f bey area -->

<!--   - here's  -->

<!-- https://www.nature.com/natsustain/about/content -->
<!-- # check in call notes -->

<!-- Ray: setting the stage: what we want is a way to assess the status of stocks and see how we're doing relative to SDG. -->

<!-- We need to discuss a sampling strategy: how can we get down to a set of priority stock. A step of expert knowledge to refine what stocks you would focus in on. -->

<!-- Regressions based approach, focusing in effort on stocks -->

<!-- Basically a paragraph laying out that we need to do this, and discussing how we might go about doing that. Equal weight to each stock vs proportional to biomass -->

<!-- add in a bit more to the discussion on guidelines -->

<!-- add in error in the index for the VOI analysis -->

<!-- Apples and apples comparisons between model and design developments. How much is the value of the effort data relative to actually improving the population model. Production model is ill-suited to some of these questions. -->

<!-- fishing effort uncertainty of the effort -->

<!-- how credible is the information -->

<!-- probably don't need to bother with the SAUP -->

<!-- stock assessments -->

<!-- the stock will collapse next year -->

<!-- nansen data -->

<!-- # Publication Requirements -->

<!-- Target Journal: [Nature Sustainability](https://www.nature.com/natsustain/about/content) -->

<!-- 3500 words main text -->

<!-- 50 references (main text) -->

<!-- 3000 words (max) methods -->

@@

The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". Meeting these targets depends in part on our ability to effectively track the status of global marine fish stocks. The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the most widely used source for tracking the global state of fisheries. As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020]. The SOFIA assessment was designed in the 1970s based on the then available data and methods, With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods, fresh approaches are needed to meet the new demand for tracking progress towards the SDG goals and supporting the research targets set out in the UNited Nations Decade on Oceans and Science plan.

Forty nine percent of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database [@ricard2012], a repository of best-available estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.

The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion. While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable, in order to increase reproducibility, transparency, and potential reliability. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts use forms of "catch-only" stock assessment models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates B/B~MSY~ and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.

Costello et al. 2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison. The super-erensemble method [@anderson2017a] used in Rosenberg et al. 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. The Pauly 2007 [@pauly2007] catch based approach finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.

@@

readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fisheries? We use a flexible surplus-production based stock assessment software package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. We show that our understanding of global fishery status is poor, and that improvements depend on an redoubled effort at global data collection.

@@

+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Data Source                                 | Short Name | Data Use                                                                 | Caveats                                                                                                         |
+=============================================+============+==========================================================================+=================================================================================================================+
| Catch data [@fao2020]                       | catches    | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Fisheries Management Index [@melnychuk2017] | FMI        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Swept Area Ratio [@amoroso2018]             | SAR        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Reconstructed effort data[@rousseau2019b]   | effort     | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors. Assumed rate of technology creep reported in individual sections |
+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+

: (\#tab:data) Broadly available data used to inform estimates of global stock status

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index (FMI) scores (a measure of management capacity), and swept area ratio (SAR) values (a measure of trawl fishing intensity) (Table.\@ref(tab:data)).

@@

For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores. We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al. 2019 [@rousseau2019] to each stock. We first used the default catch history heuristics internal to CMSY [@froese2017] to estimate stock status. We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus. Lastly, we used the reconstructed effort data from [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values. While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results (Fig.\@ref(fig:cs-plot)).

<!-- Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one. CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing. Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~. Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit. This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)). -->

```{r cs-plot, fig.cap="RAM values of B/B~MSY~ and F/F~MSY~ (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%. Each point is a stock. Black dashed line shows the 1:1 relationship."}
ex_scatter_plot
```

<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise. In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process. Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status. -->

We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status. We based this test on `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history. B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data. This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.

As a proof of concept, we also estimated B/B~MSY~ of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM. We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al. 2017 [@froese2017] (See Table.5). We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy is calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).

Overall the sraplus estimates of B/B~MSY~ resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves. This becomes clearer once we consider the performance of sraplus models fit to combinations of our broadly available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near one. This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.

Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process. Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions. For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map). We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)). Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
